
You have been given\footnote{This is adapted from Dr.\ John Ivan at UConn} the following network:
\begin{center}
\begin{tikzpicture}[auto]
  \node[cloud, align=center](A) at (0,2) {A};
  \node[cloud, align=center](B) at (0,-2) {B};
  \node[cloud, align=center](C) at (2,0){C};
  \node[cloud, align=center](D) at (6,0){D};

  \path[line](A) -- node{AD}(D);
  \path[line](A) -- node[swap]{AC}(C);
  \path[line](B) -- node{BC}(C);
  \path[line](C) -- node{CD}(D);
  \path[line](B) -- node[swap]{BD}(D);
\end{tikzpicture}
\end{center}

7000 vehicles travel from $A$ to $D$, and 5000 from $B$ to $D$ (node $C$
generates no trips). Link travel time functions are given as 
\begin{align*}
t_{AD}& = 20.0 + 0.01 V_{AD}\\
t_{AC}& = 10.0 + 0.005 V_{AC}\\
t_{CD}& = 12.0 + 0.005 V_{CD}\\
t_{BC}& = 7.25 + 0.005 V_{BC}\\
t_{BD}& = 20.0 + 0.01 V_{BD}\\
\end{align*}

\begin{enumerate}
\item{Solve for the user equilibrium (UE) link flows and travel times (HINT: write and 
solve a set of simultaneous equations that explicitly define the UE conditions). 
Demonstrate that your solution is the user equilibrium by showing through 
example that all UE conditions are satisfied.}

\item{Perform four iterations of All Or Nothing (AON) assignment on the network and 
O/D volumes. Show the link flows and travel times at the end of each iteration.}

\item{Perform an incremental assignment, using trip table increments of 25\% for each 
step. Show the link flows and travel times at the end of each incremental 
assignment.}

\item{Assign trips using the FHWA assignment heuristic. Show the link flows and 
travel times for four assignments, and the final average assignment and resulting 
travel times.}

\item{Compare the three traffic assignment heuristic approaches to the UE assignment 
and to each other. How do the resulting flow patterns differ (cite specific 
differences)? Which one comes closest to the UE flows?}

\item{State in words the general theory underlying each of the heuristic approaches. 
Which one do you prefer and why? Consider accuracy, ease of computation and 
the underlying theory.}
\end{enumerate}


All of these exercises are doable by hand. For this assignment, you may choose
to submit your answers in class on engineering calculation paper (professional
quality presentation is still expected). Students who program the solutions
(by optimizing to find the UE conditions or implementing a graph algorithm) will earn
substantial extra credit.
