Mode Choice Models
========================================================

This document contains the commands used to estimate and check mode choice models in `R`.

First, we load the data and fit it into the `mlogit` framework. We are using Bhat and Koppelman's data from the self-instructing course in discrete choice modeling developed for the FTA.
```{r loaddata, message=FALSE}
library(mlogit)
load("WorkTrips.Rdata")
# Make new id field
worktrips$IDFIELD <- paste(worktrips$HHID, worktrips$PERID, 
                           sep = "-")
# Name alternatives
alternatives <- c(" Drive Alone", " Share 2", " Share 3+", 
                  " Transit",  " Bike", " Walk")
worktrips$ALTNUM <- factor(worktrips$ALTNUM, labels = alternatives)

# make mlogit data frame
logitdata <- mlogit.data(worktrips, choice= "CHOSEN",  alt.var = "ALTNUM", 
                         chid.var = "IDFIELD", shape = "long")
```


Our first basic model contains one generic variable, `COST`, or the cost of the mode in cents.

```{r initialmodels}
base.mnl <- mlogit(CHOSEN ~ COST, data = logitdata)
summary(base.mnl)
```

What is the likelihood of this model, and how does it compare to the special market shares or null likelihoods?
```{r likelihood}
# Calculate model log-likelihood
llB <- sum(log(base.mnl$fitted.values))
llB
base.mnl$logLik

# Calculate null log-likelihood
LL0 <- 5029 * log(1/6)
LL0
# null rhosq
r20 <- 1-llB/LL0
r20

# market shares
llC <- sum(base.mnl$freq*log(prop.table(base.mnl$freq)))
llC
# constants rhosq
r2c <- 1-llB/llC
r2c
```

From this calculation, we can see that the `mlogit` package give $\rho_C^2$ as the summary default. This could be useful information. We can also estimate models with alternative-specific variables, separating the two types with the `|` ("pipe") character:
```{r ascmodel}
model2 <- mlogit(CHOSEN ~ COST | AGE, data = logitdata)
summary(model2)
```

Does this model significantly improve model likelihood? We can compare the two models with a *Likelihood Ratio Test*,
$$ -2[\ln(\mathcal{L}_1) - \ln(\mathcal{L}_2)]\sim \chi^2$$
```{r compare}
# LRT
lrstat <- -2*(base.mnl$logLik - model2$logLik)
lrstat
# find p-value
pchisq(lrstat, df=5, lower.tail=FALSE)

# Can also use function
lrtest(base.mnl, model2)
```

Does this dataset exhibit the IIA property? If it does not, then the MNL model assumptions are violated and we may need a nested logit model. We test this by estimating the model on a subset of alternatives, and comparing the full model coefficients to the reduced model coefficients with a Hausman-McFadden test. The H-M test statistic is
$$ (\beta_1 - \beta_0)'(\Sigma_0 - \Sigma_1)^{-1}(\beta_1 - \beta_0) \sim \chi^2$$
where $\Sigma$ is the variance-covariance matrix of the parameter estimates.

```{r hausmanIIA}
# estimate a model on a subset of alternatives
model2.red <- mlogit(CHOSEN ~ COST | AGE, data = logitdata, 
                 alt.subset= alternatives[1:5])

  # get coefficient names that the models have in common
  regressors <- names(coef(model2.red))[which(names(coef(model2.red)) %in% 
                                            names(coef(model2)))]
  # get relevant estimate vectors
  betas0 <- coef(model2.red)[regressors]
  betas1 <- coef(model2)[regressors]
  # get relevant variance-covariance matrices
  vcov0 <- vcov(model2.red)[regressors, regressors]
  vcov1 <- vcov(model2)[regressors, regressors]
  
  # statistic
  hmfstat <- abs((betas1-betas0) %*% MASS::ginv(vcov0-vcov1) %*% (betas1-betas0))
  hmfstat
  
  pchisq(hmfstat, length(regressors), lower.tail=FALSE)

# Again, can use function
hmftest(model2, model2.red)
```

We reject the null hypothesis that the dataset exhibits IIA, and conclude we should use a nested logit model. Let's group the car alternatives into a nest, and group the other alternatives into a different nest.

```{r nested1}
model2.nl <- mlogit(CHOSEN ~ COST + TVTT | 0, data = logitdata, 
                 nests = list(Car = alternatives[1:3], 
                              Other =alternatives[4:6]),
                 un.nest.el = FALSE)
summary(model2.nl)
```

What happens if we include alternative-specific constants in this model?
```{r nestedasc}
model3.nl <- mlogit(CHOSEN ~ COST + TVTT | 1, data = logitdata, 
                 nests = list(Car = alternatives[1:3], 
                              Other =alternatives[4:6]),
                 un.nest.el = FALSE)
summary(model3.nl)
```
The logsum parameter `iv.Car` is greater than one, implying a violation of utility theory.