Distribution Lab Solution
========================================================

We have a model with three zones. The productions and attractions vectors, as well as the cost matrix, are given below.
```{r setup}
P <- c(100, 200, 100)
A <- c(200, 50, 150)
costmatrix <- matrix(c(2,5,4,5,2,3,4,3,2), nrow=3, byrow=TRUE)
observed <- matrix(c(80,5,15,80,40,80,40,5,55), nrow=3, byrow=TRUE)
```

The script to compute the gravity model is given below.

```{r gravitymodel}
gravityModel <- function(P, A, triptime, b=1, tol = 0.01){
  
  ASTAR = A
  DIFFERENCE = Inf
  TRIPS1 = matrix(rep(0, 9), nrow=3, ncol=3)
  
  for(i in 1:100){
    if(DIFFERENCE < tol) break
    
    TRIPS <- P%o%ASTAR * triptime^-b/
      colSums(ASTAR * triptime^-b)
    
    ASTAR <- ASTAR * A/colSums(TRIPS)
    
    DIFFERENCE <- sum(abs( TRIPS1 - TRIPS ))
    TRIPS1 <- TRIPS
    #print(TRIPS)
  }
  
  return(TRIPS)
}

```

Let's find the optimum value of $b$ by minimizing an absolute error objective function.

```{r objective function}
minimumObjective <- function(b,P, A, triptime, observed){
  predicted <- gravityModel(P, A, triptime, b, tol=0.01)
  stat <- sum(abs(predicted-observed))
  return(stat)
}

saebest <- nlm(f=minimumObjective, p=1, P=P, A=A, triptime=costmatrix,
                   observed=observed)
saebest
```

According to this minimization process, the best is $b=`r saebest$estimate`$ if we judge by sum of absolute error. We can use this estimate to predict the number of trips after we make the improvements to the network:

```{r improvements}
# Original trips
gravityModel(P, A, costmatrix, saebest$estimate)

# What if improvment?
improvements <- matrix(c(2,3,4,3,2,3,4,3,2), nrow=3, byrow=TRUE)
gravityModel(P,A, improvements, saebest$estimate)
```